---
title: "R4ds Exercises"
author: "Darin Stephenson"
date: "6/1/2021"
output:
  pdf_document: default
  html_document:
    df_print: paged
---


```{r setup, include=FALSE}
require("knitr")
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir=normalizePath("..")) #set working directory to the project directory
library(conflicted)
library(tidyverse)
conflict_prefer("select", "dplyr")
conflict_prefer("lag", "dplyr")
conflict_prefer("filter", "dplyr")
```

This document is intended to help our group learn
R using the book <em>R for Data Science</em>. We can interact with R in
at least 3 distinct ways: through the console, through R scripts, and through
R Markdown documents like this one. The value of R Markdown is that you
can intersperse text and R code (and many other things) in the same
document, and then you can run the code directly in the document. It is 
also easy to produce a pdf or html output directly from an R Markdown
file using knitr.

Here is a block of R code as given in Subsection 3.2.2 of the online
version of R for Data Science.

```{r basic_plot,echo=TRUE}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))
```

Continue below to learn about data visualization from Chapter 3 of R for
Data Science. Create your own branch of the r4ds_students repository and 
push changes on your branch to GitHub regularly (at least each day).